Delayed effects of climate on vital rates leads to demographic divergence in Amazonian forest fragments

### 2021-05-25 doc/paper.Rmd
add in final citation

Third, it may be that the delayed demographic effects we observed are indirectly mediated by the effect of SPEI on other species rather than the direct effects on individual physiology [@eversLaggedDormantSeason2021].
For example, topical trees may not die until three or more years after a drought [@crileyYearProductionHigh1994].
When they finally do, the resulting leaf drop [@janssenDroughtEffectsLeaf2021] and treefalls allow for light penetration to the forest understory [@canham1990; @leitoldNinoDroughtIncreased2018], triggering a boom in the growth and flowering of understory plants [@brunaDemographicEffectsHabitat2005].
Similar delayed changes in the local environment could also influence the foraging behavior of a plant's pollinators [@brunaHeliconiaAcuminataReproductive2004; @stouffer1996], seed dispersers [@uriarteDisentanglingDriversReduced2011], or herbivores [@scottDifferingNonlinearLagged2021].
While more work is needed to explain why the (delayed) effects of SPEI on *H. acuminata* survival and growth are greater in fragments than forest interiors, one hypothesis motivated by recent intriguing results from other systems **(CITATIONS)** <!--#  --> is that the greater litterfall on edges [@vasconcelos2004] may be altering the abundance of pathogens or mycorrhizae.

>>>>

Third, it may be that the delayed demographic effects we observed are indirectly mediated by the effect of SPEI on other species rather than the direct effects on individual physiology [@eversLaggedDormantSeason2021].
For example, topical trees may not die until three or more years after a drought [@crileyYearProductionHigh1994].
When they finally do, the resulting leaf drop [@janssenDroughtEffectsLeaf2021] and treefalls allow for light penetration to the forest understory [@canham1990; @leitoldNinoDroughtIncreased2018], triggering a boom in the growth and flowering of understory plants [@brunaDemographicEffectsHabitat2005].
Similar delayed changes in the local environment could also influence the foraging behavior of a plant's pollinators [@brunaHeliconiaAcuminataReproductive2004; @stouffer1996], seed dispersers [@uriarteDisentanglingDriversReduced2011], or herbivores [@scottDifferingNonlinearLagged2021].
While more work is needed to explain why the (delayed) effects of SPEI on *H. acuminata* survival and growth are greater in fragments than forest interiors, one hypothesis, motivated by recent intriguing results from other systems [@sapsfordChickenEggWhich2017], is that the greater litterfall on edges [@vasconcelos2004] may be altering the abundance of pathogens or mycorrhizae.
